---
title: "Lecture-A Examples"
author: "Christopher Prener, Ph.D."
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output: 
  github_document: default
  html_notebook: default 
---

## Introduction
This notebook provides additional examples of data cleaning plus new material on table joins and exporting data.

## Dependencies
This notebook requires a number of packages:

```{r load-packages}
# tidyverse packages
library(readr)       # read/write tabular data

# spatial packages
library(mapview)     # preview spatial data
library(sf)          # spatial data tools

# other packages
library(here)        # file path management
```

## Load Data
This notebook requires one set of data:

```{r load-data}
sluPlaces <- read_csv(here("data", "example-data", "sluPlaces.csv"))
```

## Data Cleaning Notes
Make sure your `x` and `y` coordinate variables are numeric or double:

```{r class}
sluPlaces2 <- mutate(sluPlaces, lng = as.numeric(lng))
sluPlaces2 <- mutate(sluPlaces2, lat = as.numeric(lat))
```

If they are not, use `mutate()` with `as.numeric()` to convert them.

## Project Data
First, we want to convert these data to from a `tibble` to an `sf` object with `st_as_sf()`. We use the `lng` variable as our `x` variable and `lat` as our `y` variable, and use `4269` for our `crs` argument since these data are in decimal degrees:

```{r project}
sluPlaces_sf <- st_as_sf(sluPlaces, coords = c("lng", "lat"), crs = 4269)
```

Next, we want to confirm that this worked:

```{r preview}
mapview::mapview(sluPlaces_sf)
```

Excellent!

## Transform Our Projection
We've already used `st_transform()`, but now can do so with purpose. For example, to convert our data to State Plane (feet):

```{r convert-to-state-plane}
sluPlaces_sf <- st_transform(sluPlaces_sf, crs = 102696)
```

We'll confirm that this is the right projection:

```{r}
st_crs(sluPlaces_sf)
```

## Write Data
Finally, we'll write our data:

```{r write-data}
st_write(sluPlaces_sf, here("data", "example-data", "clean-data", "sluPlaces.shp"),
         delete_dsn = TRUE)

st_write(sluPlaces_sf, here("data", "example-data", "clean-data", "sluPaces.geojson"),
         delete_dsn = TRUE)
```

If we wanted to read these back in:

```{r read-back-in}
sluPlaces_sf_2 <- st_read(here("data", "example-data", "clean-data", "sluPaces.geojson"),
                          crs = 102696)
```



