---
title: "Lab-A Replication"
author: "Christopher Prener, Ph.D."
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output: 
  github_document: default
  html_notebook: default 
---

## Introduction
This notebook replicates the results of Lab-A.

## Dependencies
This notebook requires the following packages:

```{r load-packages}
# tidyverse packages
library(dplyr)    # data manipulation
library(ggplot2)  # mapping
library(readr)    # import tabular data

# spatial packages
library(mapview)  # interactive map preview
library(tmap)     # mapping
library(sf)       # methods for spatial data

# other packages
library(here)     # file path management
```

## Load Data
This notebook requires three data sets:

```{r load-data}
# spatial 
county_bnd <-st_read(here("data","US_BOUNDARY_Counties","US_BOUNDARY_Counties.shp"), stringsAsFactors = FALSE)

# tabular
capitals <- read_csv(here("data", "stateCapitals.csv"))
health_ins <- read_csv(here("data","US_HEALTH_noIns.csv"))
```

## Part 1
### Question 1
#### A
First, we'll project the state capitals data. THe data are in decimal degrees, so we'll use the geogrpahic coordinate system NAD 1983 to project them. The `crs` value for that projection is 4269

```{r p1-q1-a}
# project
capitals <- st_as_sf(capitals, coords = c("longitude", "latitude"), crs = 4269)

# check
mapview(capitals)
```

#### B
We have projected data! We'll go ahead and save these data, and then remove Alaska and Hawaii:

```{r p1-q1-a-2}
# save data
st_write(capitals, here("data", "cleanData", "US_CAPITALS_State", "US_CAPITALS_State.shp"),
         delete_dsn = TRUE)

# subset
capitals <- filter(capitals, name %in% c("Alaska", "Hawaii") == FALSE)
```

#### C
Next, we need to combine our county data in `county_bnd` with the `health_ins` data. First, we want to join these data. `GEOID` appears to be a common variable among them, but it is character in one data.frame and numeric in the other. To fix this, we can `mutate` one class to match the other.
```{r p1-q1-c}
health_ins <- mutate(health_ins, GEOID = as.numeric(GEOID))
```

Now we can table join as usual.

```{r p1-q1-c-2}
cnty_health <- left_join(county_bnd, health_ins, by = "GEOID")
```

#### Part D
We want to subset our data to remove the observations with missing data. Since we know that $-1$ denotes missing data, we'll `filter` for observations where `noIns` is $>=0$

```{r p1-q1-d}
cnty_health <- filter(cnty_health, noIns >= 0)
```

#### Part E
Next, we'll repeat the same subsetting that we did with the capitals:

```{r p1-q1-e}
# subset
cnty_health <- filter(cnty_health, state %in% c("Alaska", "Hawaii") == FALSE)
```

We could also use the `STATEFP` variable:

```r
cnty_health <- filter(cnty_health, STATEFP %in% c("02", "15") == FALSE)
```

Both approaches work equally well in this case.

#### Part F
Finally, we'll export our data:

```{r p1-q1-f}
st_write(cnty_health, here("data", "cleanData", "US_HEALTH_noIns", "US_HEALTH_noIns.shp"),
         delete_dsn = TRUE)
```

## Part 2
### Question 2
Next, we'll apply the Lambert project coordinate system to both of our data objects:

```{r p2-q2}
### insurance data
cnty_health <- st_transform(cnty_health, crs = 102004)

### capitals
capitals <- st_transform(capitals, crs = 102004)
```

Now our data are ready for mapping.

### Questions 3-5, `ggplot2`
First, we'll make our plot using `ggplot2`. I'll use the `lwd` argument to diminish the lines around each of the counties:

```{r p2-q3-5-ggplot2}
ggplot() +
  geom_sf(data = cnty_health, mapping = aes(fill = noIns), lwd = .15) +
  geom_sf(data = capitals) +
  scale_fill_distiller(palette = "YlGn", trans = "reverse", name = "Percent") +
  labs(
    title = "Percent of Residents without Health Insurance",
    subtitle = "By U.S. County",
    caption = "Map by Christopher Prener, Ph.D.\nData via the CDC"
  ) +
  theme_void() -> map1

map1
```

### Questions 3-5, `tmap`
First, we'll make our plot using `ggplot2`. I'll use the `lwd` argument to diminish the lines around each of the counties:

```{r p2-q3-5-tmap}
tm_shape(cnty_health) +
  tm_polygons(col = "noIns",
              palette = "YlGn",
              style = "jenks",
              title = "Percent", 
              lwd = .25) +
  tm_shape(capitals) +
    tm_symbols(
      col = "#000000",
      size = 0.25) +
  tm_layout(
    title = "Percent of\nResidents without\nHealth Insurance",
    frame = FALSE,
    legend.outside = TRUE,
    legend.position = c("right", "bottom")
  ) -> map2

map2
```

### Question 6, `ggplot2`
Next, we'll save our `ggplot2` map:

```{r p2-q6-ggplot2}
ggsave(here("results", "ggplot2_map.png"), map1, dpi = 500)
```

### Question 6, `tamp`
Finally, we'll save our `tmap` map:

```{r p2-q6-tmap}
tmap_save(tm = map2, filename = here("results", "tmap_map.png"), dpi = 500)
```
